Data Cleaning

```{r setup}

#loading in packages
library(titanic)
library(dplyr)
library(janitor)
library(ggplot2)
library(tidyr)
library(Hmisc)
library(stringr)
```

Link to Dataset: <https://www.kaggle.com/datasets/jainaru/global-ecological-footprint-2023>

-   The Kaggle view is very helpful and easy to read

```{r}

#loading in the dataset
fileName = list.files("data")
data <- read.csv(paste0("data/",fileName))

#looking at part of the data
data
```

```{r}

#clean names
data <- clean_names(data, case="snake")


data <- rename(data, 
life_expectancy = life_exectancy, #spell expectancy correctly
sdgi = sd_gi, #get rid of the underscore
bu_land_footprint = built_up_land, #differentiate and shorthand the "built up lands"
bu_land_biocap = built_up_land_1, 
eco_difference = ecological_deficit_or_reserve, 
total_footprint = total_ecological_footprint_consumption, 
num_earth = number_of_earths_required, 
num_countries = number_of_countries_required)
#the last 4 just make them easier to work with
```

```{r}
#filling in empty strings as nulls
data[data == ""] <- NA
data[data == " "] <- NA

#making population numeric data
data$population_millions <- gsub(",", "", data$population_millions)
data = mutate(data, population_millions = as.numeric(population_millions))

#making gdp numeric
data$per_capita_gdp <- gsub(",", "", data$per_capita_gdp)
data$per_capita_gdp <- gsub("\\$", "", data$per_capita_gdp)
data = mutate(data, per_capita_gdp = as.numeric(per_capita_gdp))


head(data)
```

No need to remove_empty since there are no entry points

```{r}

#dimensions of data
cat("The Dimensions are:",nrow(data), "x", ncol(data))
```

```{r}

#pandas .info()
str(data) #look at console
```

EDA on the footprint features

```{r}

#select the features ending with "footprint"
footprint_features <- select(data, "country", contains("footprint"))
print(footprint_features)
summary(footprint_features) #look at the console for this one
#total eco footprint is the sum of the other "footprints"
```

Countries vs. Their More Composite Features

```{r}
#country versus its total eco footprint
select(data, "country", "total_footprint")
```

```{r}

#country versus its eco difference
select(data, "country", "eco_difference")
```

```{r}

#HDI vs. eco difference
select(data, "country", "hdi", "eco_difference")
```

Making New Features

-   thanks Stephen for the idea of using the combo of eco deficit and num earth (idk if imma do this)

```{r}

#eco difference as a binary variable
featured_data <- mutate(data, is_sustainable = eco_difference>0)
head(featured_data)

#a country with a positive eco_difference is sustainable because they used less resources than their country can provide

#plot the count of each
ggplot(featured_data, mapping = aes(x=is_sustainable)) + geom_bar() + labs(title = "Number of Sustainable Countries", x="Is Sustainable")
```

```{r}
#sdgi, life expectancy, per capita gdp, population - features not directly associated with footprint and sustainability

other_and_sustainable_data = select(featured_data, "country", "sdgi", "hdi", "life_expectancy", "per_capita_gdp", "population_millions", "eco_difference", "is_sustainable")
other_and_sustainable_data = filter(other_and_sustainable_data, !is.na(is_sustainable))

other_and_sustainable_data
```

Possible Research Question: Do \[these\] factors correlate on whether a country is self sustainable, and if they do, how do they correlate?

EDA and Plotting of these Features

```{r}
#repeat of previous graph
#chart of number of self sustainable countries
ggplot(featured_data, mapping = aes(x=is_sustainable)) + geom_bar(fill='#0081c2') + labs(title = "Number of Sustainable Countries", x="Is Sustainable")
```

```{r}
#look at console for summary of data
summary(other_and_sustainable_data)
```

```{r}
#the max for eco difference is 84 and I want to know which country that is
arrange(other_and_sustainable_data, desc(eco_difference))

#follow up: why do French Guinana, Suriname, and Guyana (all next to each other) have such a positive eco difference
```

```{r}
#plotting population vs. eco difference, not using the binary variable here because it makes the graphing much more confusing

ggplot(other_and_sustainable_data, mapping=aes(x=population_millions, y=eco_difference, color=population_millions)) + geom_point() + scale_x_log10(breaks=c(1e-01, 1e+00, 1e+01, 1e+02, 1e+03), labels=c("0.1", "1", "10", "100", "1000")) + ylim(-20, 20) + labs(title="Population vs. Eco Difference", x="Population (in millions, on a log scale)", y="Ecological Difference")

#the y limit does eliminate the guyana, suriname, and french guinana outliers
#note-population_millions is on a log scale

#there isn't an obvious trend between population and eco_difference
#however, it might become more clear in the model when using the binary target variable
```
