---
title: "The Correlation Between a Country’s Financial Status and Its Sustainability"
execute:
  echo: FALSE
  warning: FALSE
  error: FALSE
---

by Pranav Sant

## Background

The average global temperature has skyrocketed over the years, caused by excess carbon emissions and waste. This is primarily caused by a global ecological footprint that significantly exceeds our bio-capacity, causing us to deplete our resources as well as release waste and carbon into the environment. Because some countries contribute to this unsustainable behavior more than others, it is important to explore any possible trends or reasons that may cause this.

Two important metrics that we will use repeatedly to measure sustainability of countries are "Number of Earths Required" and "Ecological Reserve". The number of earths required represents the number of earths we would need to be sustainable if everyone lived like the average person in this country. The ecological reserve is the resources/capacity a country has minus how much it uses, measured in hectares per person.

## Results

### Initial Data Exploration and Plotting

```{r setup}
library(dplyr)
library(tidyr)
library(janitor)
library(ggplot2)
library(caTools)
library(FNN)

clean_data <- read.csv("C:/Users/pranav/Documents/Coding/DSRP-2024-George/PranavProject/Data/clean_data.csv")
```

```{r}
ggplot(data = clean_data, aes(x = per_capita_gdp, y = ecological_deficit_or_reserve, color = hdi)) +
  geom_point() +
  labs(title = "GDP Per Capita vs Ecological Reserve by Human Development Index",
       x = "GDP Per Capita ($)",
       y = "Eco Reserve (Hectares per Person)") +
  geom_smooth(color = "red") + ylim(-15, 15)
```

**Figure 1**. The correlation between GDP Per Capita and Ecological Reserve with Human Development Index

This graph shows that as GDP Per Capita increases, the ecological reserve tends to decrease, showing the countries with a greater per capita GDP are generally more unsustainable.

```{r}
ggplot(data = clean_data, aes(x = per_capita_gdp, y = number_of_earths_required, color = hdi)) +
  geom_point() +
  labs(title = "GDP Per Capita vs Number of Earths Required by HDI",
       x = "GDP Per Capita",
       y = "Earths Required") +
  geom_smooth(color = "red")
```

**Figure 2**. The correlation between GDP Per Capita and Number of Earths Required with Human Development Index

This graph also supports a similar conclusion to Figure 1. Countries with a higher per capita GDP have an average lifestyle that requires a greater amount of earths, pointing to unsustainability.

```{r}
grouped_data <- clean_data %>%
  mutate(group = case_when(
    gdp < 27885 ~ 1,
    gdp > 27885 & gdp < 100136.2 ~ 2,
    gdp > 100136.2 & gdp < 438837.8 ~ 3,
    gdp > 438837.8 ~ 4
  ))

group1 <- subset(grouped_data, group == 1)
group2 <- subset(grouped_data, group == 2)
group3 <- subset(grouped_data, group == 3)
group4 <- subset(grouped_data, group == 4)

grouped_avg <- aggregate(grouped_data$number_of_earths_required, list(grouped_data$group), FUN = mean) 

ggplot(data = grouped_avg, aes(x = Group.1, y = x)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Number of Earths Required by GDP Group", x = "Group", y = "Earths Required")
```

**Figure 3**. The correlation between GDP and Earths Required when GDP is split into groups by quartile. The groups are in order of increasing GDP, with Group 1 having the least GDP and Group 4 having the greatest GDP.

While Figure 1 and Figure 2 use GDP per capita as a their metric, Figure 3 uses GDP. The data had to be grouped due to some outlier data points, but this graph also shows that countries with stronger economies tend to require more earths.

```{r}

reordered_income_group <- factor(clean_data$income_group, levels=c("LI", "LM", "UM", "HI"))


ggplot(data = clean_data, aes(x = reordered_income_group, y = number_of_earths_required)) +
  geom_boxplot() +
  labs(title = "Earths Required by Income Group",
       x = "Income Group",
       y = "Earths Required")
```

**Figure 4**. Box Plots of Various Income Groups and their respective Earths Required

While figures until now have relied on GDP per capita and GDP metrics, this graph is based on four Income Groups that were defined through the original dataset.

LI = Low Income, LM = Lower Middle Income, UM = Upper Middle Income, HI = High Income, while the NA category contains countries who were not assigned a group.

This collection of box plots supports the same conclusions as Figures 1-3, showing that countries with superior financial status tend to be more unsustainable.

```{r}
ggplot(data = clean_data, aes(x = reordered_income_group, y = ecological_deficit_or_reserve)) +
  geom_boxplot() +
  labs(title = "Ecological Reserve by Income Group",
       x = "Income Group",
       y = "Ecological Reserve") + ylim(-15, 15)
```

**Figure 5**. Box Plots of Various Income Groups and their respective Ecological Reserve

Just like Figure 4, this graph uses income groups to demonstrate the more prominent overuse of resources by countries with a higher income. Figures 1-5 have so far supported that the Earths Required and Ecological Reserve are highly similar in terms of how they are reflected by the financial status of countries.

```{r}
ggplot(data = clean_data, aes(x = region, y = ecological_deficit_or_reserve)) +
  geom_boxplot() +
  labs(title = "Ecological Reserve by Region",
       x = "Region",
       y = "Ecological Reserve") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) + ylim(-15, 15)
```

**Figure 6**. Box Plots of Various Regions and their respective Ecological Reserve

Instead of directly focusing on finances, Figure 6 examines the correlation between countries' regions and their ecological reserve through a collection of boxplots for each region. The regions that tend towards being the most unsustainable are Central Asia, Europe, and North America.

```{r}
ggplot(data = clean_data, aes(x = region, y = number_of_earths_required)) +
  geom_boxplot() +
  labs(title = "Earths required by Region",
       x = "Region",
       y = "Earths Required") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

**Figure 7**. Box Plots of Various Regions and their respective Earths Required

This graph is similar to Figure 6 in that it examines sustainability by the region metric. The most unsustainable regions by Earths Required are North America and Europe. When comparing Figure 6 and Figure 7, we can see one of the first times that the metrics of Earths Required and Ecological Reserve present a slighlty different conclusion, as the regions that are more sustainable than others changes depending on which metric we chose.

### Hypothesis Testing

#### T-testing

The Human Development Index measures basic human development on a scale of 0-1.0, taking into account life expectancy, standard of living, and expected years of schooling. This is generally considered an excellent metric of a country's financial and developmental status, and thus we can use it as a reliable value to test correlation with sustainability.

We split the data into two groups, one called "Low HDI" with an HDI of less than 0.5, and the other called "High HDI", with HDI greater than or equal to 0.5.

Then, we ran two T-tests to find out if the HDI had a significant correlation with the Earths Required and the Ecological Reserve. For both tests, we set the threshold for significance at 0.05, as we wanted a 95% confidence interval.

```{r}

high_hdi <- subset(clean_data, hdi >= 0.5)
low_hdi <- subset(clean_data, hdi < 0.5)
```

The first test was of the Earths Required. Here, the null hypothesis was: the low HDI and the high HDI have no significant difference in the number of earths required. See Figure 2 for the plot that shows the trend apparent in the data.

```{r}
t.test(high_hdi$number_of_earths_required, low_hdi$number_of_earths_required, mu = mean(clean_data$number_of_earths_required, na.rm = TRUE))
```

This test yielded a p-value that was far less than 0.05, and thus we determined that the high and low HDI groups have a significant difference in Earths Required, allowing us to reject the null hypothesis.

For the second T-test, we used Ecological Reserve. In this case, the null hypothesis was: the low HDI and the high HDI have no significant difference in ecological reserve. See Figure 1 for the plot that shows the trend apparent in the data.

```{r}
t.test(high_hdi$ecological_deficit_or_reserve, low_hdi$ecological_deficit_or_reserve, mu = mean(clean_data$ecological_deficit_or_reserve, na.rm = TRUE))

```

This test yielded a p-value that was greater than 0.05, and thus we the high and low HDI groups do not have a significant difference in Ecological Reserve, meaning we fail to reject the null hypothesis.

#### ANOVA testing

We used ANOVA testing to see if there was a significant difference in the mean of the Earths Required and Ecological Reserve between Income Groups. Once again, we used a confidence interval of 95%, meaning we need a p-value less than 0.05 for significance.

First, we looked at Earths Required. The null hypothesis here is: there is no significant difference in the earths required between the different income groups.

```{r}
earths_aov <- aov(data = clean_data, number_of_earths_required ~ income_group)
summary(earths_aov)
```

The ANOVA test yields a p-value far less than 0.05, and thus we can confidently reject the null hypothesis, proving that income group is correlated with earths required.

Next, we looked at Ecological Reserve. The null hypothesis here is: there is no significant difference in the ecological reserve between the different income groups

```{r}
reserve_aov <- aov(data = clean_data, ecological_deficit_or_reserve ~ income_group)
summary(reserve_aov)
```

The result here also has a p-value less than 0.05, so we can also reject this null hypothesis. Through this, we know that unlike HDI, income group has a strong correlation with sustainability using both metrics, not just Earths Required.

We conducted similar testing for the region, seeing if there was a significant difference in the sustainability of different areas of the world. The confidence interval remained at 95%. The results were as follows.

For Region vs Earths Required, the Null Hypothesis was: There is no significant difference in the earths required of any two regions.

```{r}
region_earths_aov <- aov(data = clean_data, number_of_earths_required ~ region)
summary(region_earths_aov)
```

The p-value is less than 0.05, so we reject the null hypothesis.

For Region vs Ecological Reserve, the Null Hypothesis was: There is no significant difference in the ecological reserve of any two regions.

```{r}
region_reserve_aov <- aov(data = clean_data, ecological_deficit_or_reserve ~ region)

summary(region_reserve_aov)
```

The p-value is less than 0.05, so we once again reject the null hypothesis.

Regarding the region metric, both tests yielded significant results, showing that there is a significant difference in sustainability by region using both earths required and ecological reserve.

### Machine Learning Models

To see how accurately we could predict various metrics using the correlation between financial status and sustainability, we used the machine learning models Logistic Regression and K-Nearest Neighbors and tested their accuracy.

#### Logistic Regression

We used the logistic regression to predict the earths required using the metrics of GDP per capita and HDI.

```{r}
lr_split <- sample.split(clean_data$number_of_earths_required, SplitRatio = 0.8)
lr_train_data <- subset(clean_data, lr_split == TRUE)
lr_test_data <- subset(clean_data, lr_split == FALSE)

lr_model <- lm(number_of_earths_required ~ hdi + per_capita_gdp, data = lr_train_data)

lr_pred <- predict(lr_model, newdata = lr_test_data)


ggplot(data = lr_test_data, aes(x = number_of_earths_required, y = lr_pred)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "solid", color = "red", lwd = 1) +
  labs(title = "Real vs Predicted Values in Linear Regression Model",
       x = "True Values",
       y = "Predicted Values")

lr_mse <- mean((lr_test_data$number_of_earths_required - lr_pred)^2, na.rm = TRUE)
lr_r2 <- summary(lr_model)$r.squared

cat("R squared value:", lr_r2)
cat("\n")
cat("Mean Squared Error:", lr_mse)

```

**Figure 8**. A graph of true vs predicted values of Earths Required from the Logistic Regression.

Depending on the split, the R squared value for this model can range from 0.5 to 0.9 and the Mean Squared Error can range from 0.35 to 1.3. Overall, this model shows mediocre accuracy. However, it provest that there is a trend in the data.

#### K-Nearest Neighbors

We used a K-nearest neighbors model to predict the income group using Earths Required and Ecological Reserve.

```{r}
knn_data <- clean_data[, c('income_group', 'ecological_deficit_or_reserve', 'number_of_earths_required')]

knn_data <- na.omit(knn_data)

knn_split <- sample.split(knn_data$income_group, SplitRatio = 0.8)
knn_train_data <- subset(knn_data, knn_split == TRUE)
knn_test_data <- subset(knn_data, knn_split == FALSE)


knn_model <- knn(train <- knn_train_data[, c('number_of_earths_required', 'ecological_deficit_or_reserve')],
                 test <- knn_test_data[, c('number_of_earths_required', 'ecological_deficit_or_reserve')],
                 cl = knn_train_data$income_group)


pred <- knn_model

actual <- knn_test_data$income_group

cm <- table(actual, pred)
cat("Confusion Matrix:\n")
cm

cat("Accuracy:", sum(diag(cm)) / sum(cm))
```

Depending on the split, the Confusion Matrix accuracy can range from 0.4 to 0.7. Overall, this model also shows mediocre accuracy, but it also proves that there is a real trend.

## Discussion

From Figures 1-5, there is an apparent trend that shows countries with a greater financial status (GDP per capita, GDP, Income Group, HDI) consistently being more unsustainable. They generally require more earths and have a lower ecological reserve. This observation indicates that there is a clear difference in the lifestyle and resource usage of these countries that leads to their lack of sustainability.

Regarding region, based on the box plots in Figure 6 and Figure 7, areas that are typically considered "wealthy" by traditional standards have greater unsustainability, supporting the idea that a greater financial status leads to environmentally unfriendly practices.

Through the Statistical Testing, we can see that the Human Development Index metric is not as reflective of sustainability as the income group, as it has no significant correlation to the ecological reserve. All of the other tests came out to be significant, showing that there is a major trend and thus, it is possible to initiate change to better our practices.

While there were significant results, more research still needs to be done, as the dataset had some limitations. Not all values were filled, and so some countries were excluded from the testing. Additionally, although we know that wealthier countries consume more, we still need to look deeper into what resources exactly they are consuming and the causes of this. Looking deeper into the problem will thus allow us to combat the threat of climate change.

The ideal solution seems to be for people within many of the countries with a greater financial status to change their lifestyle, putting a greater emphasis on sustainability; however, this is impractical as people will have to sacrifice comfort and convenience to do this, which many will be unwilling to do.

One way we could combat our environmental problems is to advocate for change, pressing the government to use and encourage more sustainable practices.

Although we still lack a definite plan to avert the pressing issue of climate change, it is still important to investigate potential causes for it, and thus, this research serves as one of the first steps to combat this problem.

## Code and Data Availability

The dataset that we used is available at <https://www.kaggle.com/datasets/jainaru/global-ecological-footprint-2023> and the GitHub repository can be found here: <https://github.com/the-codingschool/DSRP-2024-George/tree/dev-Pranav>

## Acknowledgements

I would like to acknowledge George Zhou for guiding me through this research project, helping me properly understand the data and providing me with questions to explore along my research journey. I would also like to thank Shruti M and Sarah Parker, who helped teach me the R language and answered countless programming questions throughout my research process. Also, thanks to my lab group members for supporting me throughout.
